---
title: "Study 5 - Apoptosis and other immune biomarkers predict influenza vaccine responsiveness (PMID: 23591775)"
author: "Mikkel Niklas Rasmussen"
date: "5/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Applying the 23 differentially expressed genes between the non-responder and responder at day 0 presented in study by Bartholomeus et al. (PMID: 30205979) on the CMI-PD dataset

```{r load libaries, echo=FALSE}
setwd(dirname(rstudioapi::getSourceEditorContext()$path))

suppressMessages(library('data.table'))
suppressMessages(library('dplyr'))
suppressMessages(library('plyr'))
suppressMessages(library('tidyr'))
suppressMessages(library('Biobase'))
suppressMessages(library('DESeq2'))
suppressMessages(library('pROC'))
suppressMessages(library("ggplot2"))
suppressMessages(library("RCurl"))
suppressMessages(library("readr"))
suppressMessages(library("biomaRt"))
suppressMessages(library('klaR'))
suppressMessages(library("tibble"))

# Load functions for evaluating gene signatures
source('../R/modelEvaluate.R')

```

#### Load the CMI-PB datasets

```{r load data}

DATA_DIR_raw <- "../Data/raw"
DATA_DIR_full <- "../Data/processed/full"

# Info table (meta data)
meta.2020 <- fread(file.path(DATA_DIR_full, "meta.2020.pivoted.tsv"))
meta.2021 <- fread(file.path(DATA_DIR_full, "meta.2021.pivoted.tsv"))
specimen.table.2020 <- fread(file.path(DATA_DIR_raw, "2020LD_specimen.csv"))
specimen.table.2021 <- fread(file.path(DATA_DIR_raw, "2021BD_specimen.csv"))

# Antibody titer data
ab.titer.2020 <- fread(file.path(DATA_DIR_raw, "2020LD_ab_titer.csv"))
ab.titer.2021.day0 <- fread(file.path(DATA_DIR_raw, "2021BD_ab_titer.csv"))
ab.titer.2021.day14 <- fread(file.path(DATA_DIR_raw, "../2021_ab_titer_day14.csv"))

```

## Step 1: Preprocess CMI-PB 2020 antibody titer data

```{r}

# Normalized antibody titer data by baseline median
abt.2020.all <- ab.titer.2020 %>% 
   filter(antigen %in% c('PT', 'PRN', 'FHA', 'FIM2/3')) %>% 
   unite(isotype_antigen, c("isotype", "antigen"), sep = "-") %>% 
   inner_join(., specimen.table.2020 %>% dplyr::select(specimen_id, subject_id, 
                                                  planned_day_relative_to_boost), 
              by='specimen_id') %>% 
   filter(planned_day_relative_to_boost %in% c(0, 1, 3, 7, 14, 30)) %>% 
   dplyr::rename(timepoint = planned_day_relative_to_boost) %>% 
   tibble()

# Select the subject that have data day 0 and 14
subjects.day0 <- unique(abt.2020.all[abt.2020.all$timepoint == 0, ]$subject_id)
subjects.day14 <- unique(abt.2020.all[abt.2020.all$timepoint == 14, ]$subject_id)
ids <- intersect(subjects.day0, subjects.day14)

# Subset antibody titer
abt.2020.subset <- abt.2020.all[abt.2020.all$subject_id %in% ids, ]
abt.2020.day0 <- abt.2020.subset[abt.2020.subset$timepoint == 0, ]
abt.2020.day14 <- abt.2020.subset[abt.2020.subset$timepoint == 14, ]

```

## Step 2: Evaluation age of subjects as predictor of the antibody response

```{r}

# Isotype-antigens which are considered the most important
selected.iso.ag <- c("IgG-PT", "IgG-FHA",   "IgG-PRN", "IgG1-PT",
                     "IgG1-FHA", "IgG4-PT", "IgG4-FHA")

################ Evaluating age as a predictor ###############
# Calculate the age of the subjects
df.age.2020.pred <- mlModelPredict(meta.data=meta.2020, 
                                   study='Furman_2013',
                                   post.ab.data=abt.2020.day14,
                                   pre.ab.data=abt.2020.day0,
                                   filter.ab.ag=selected.iso.ag)

# Evaluate age as a predictor
Furman.2013.Age.2020.results <- evaluateModel(df.prediction=df.age.2020.pred,
                                              N.perm = 10000, direction = ">",
                                              score.label = 'Age',
                                              fn1='Age_vs_day14_heatmap',
                                              fn2='Age_vs_day14_correlation',
                                              fn3='Age_vs_FC_heatmap',
                                              fn4='Age_vs_FC_correlation',
                                              fn5='Age_vs_day14_AUC',
                                              fn6='Age_vs_day14_boxplot',
                                              fn7='Age_vs_FC_AUC',
                                              fn8='Age_vs_FC_boxplot',
                                              path='Results/2020/Age/')

# Save results
save(Furman.2013.Age.2020.results, 
     file = "Results/2020/Age/Furman.2013.Age.2020.results.RData")

```

## Step 3: Pre-process CMI-PB 2021 antibody titer data

```{r}

# Normalized antibody titer data by baseline median
ab.titer.day0.2021.all <- ab.titer.2021.day0 %>% 
   filter(antigen %in% c('PT', 'PRN', 'FHA', 'FIM2/3')) %>% 
   unite(isotype_antigen, c("isotype", "antigen"), sep = "-") %>% 
   inner_join(., specimen.table.2021 %>% dplyr::select(specimen_id, subject_id, 
                                                  planned_day_relative_to_boost), 
              by='specimen_id') %>% 
   dplyr::rename(timepoint = planned_day_relative_to_boost) %>% 
   tibble()

ab.titer.subject.ids <- unique(ab.titer.day0.2021.all$subject_id)
abt.2021.day0 <- ab.titer.day0.2021.all[ab.titer.day0.2021.all$timepoint == 0, ]
abt.2021.day0 <- abt.2021.day0 %>% 
   filter(isotype_antigen %in% selected.iso.ag)

# Change format of the day 14 antibody titers
abt.2021.day14 <- ab.titer.2021.day14 %>% 
   separate(task, into=c("task", "antigen"), sep = "\\) ") %>% 
   dplyr::mutate(isotype_antigen = recode(antigen, 
                                          `Pertussis Toxin` = "IgG-PT", 
                                          `FHA` = 'IgG-FHA', 
                                          `Pertactin` =  'IgG-PRN',
                                          `IgG1 - Pertussis toxin` = "IgG1-PT",
                                          `IgG1 - FHA` = "IgG1-FHA",
                                          `IgG4 - Pertussis toxin` = "IgG4-PT",
                                          `IgG4 - FHA` = "IgG4-FHA"))


```

## Step 4: Make predictions for CMI-PB 2021 data 

```{r}

# Subset meta data based on subjects with antibody titers
meta.2021.subset <- meta.2021[meta.2021$subject_id %in% ab.titer.subject.ids, ]

# Calculate the age of the subjects
df.age.pred.2021 <- mlModelPredict(meta.data=meta.2021.subset, 
                                   study='Furman_2013')

# Read submisson format
submission.format <- fread("../../Data/2021_cmipb_challenge/2021_submission_format.tsv") %>% 
   dplyr::rename(subject_id = Subject_ID)

# Age predictions
Furman.2013.Age.2021.pred <- submission.format %>% 
   left_join(., df.age.pred.2021, by="subject_id")
Furman.2013.Age.2021.pred[, c(5:11)] <- Furman.2013.Age.2021.pred$rank.predictor
Furman.2013.Age.2021.pred <- Furman.2013.Age.2021.pred[,c(1:18)] %>% 
   dplyr::rename(Subject_ID = subject_id)
# Save predictions
write.table(Furman.2013.Age.2021.pred, 
            file = "Results/2021_predictions/furman_2013_age_CMI-PB_2021.tsv",
            quote=FALSE, sep='\t', row.names = FALSE)

```

## Step 9: Evaluate the DLDA model using different number of cell populations on the CMI-PB 2021 dataset

```{r model evaluation}

# Calculate the age of the subjects
df.age.2021.pred <- mlModelPredict(meta.data=meta.2021, 
                                   study='Furman_2013',
                                   post.ab.data=abt.2021.day14,
                                   pre.ab.data=abt.2021.day0,
                                   filter.ab.ag=selected.iso.ag)

# Evaluate age as a predictor
Furman.2013.Age.2021.results <- evaluateModel(df.prediction=df.age.2021.pred,
                                              N.perm = 10000, direction = ">",
                                              score.label = 'Age',
                                              fn1='Age_vs_day14_heatmap',
                                              fn2='Age_vs_day14_correlation',
                                              fn3='Age_vs_FC_heatmap',
                                              fn4='Age_vs_FC_correlation',
                                              fn5='Age_vs_day14_AUC',
                                              fn6='Age_vs_day14_boxplot',
                                              fn7='Age_vs_FC_AUC',
                                              fn8='Age_vs_FC_boxplot',
                                              path='Results/2021/Age/')

# Save results
save(Furman.2013.Age.2021.results, 
     file = "Results/2021/Age/Furman.2013.Age.2021.results.RData")

```
