---
title: "Study 2 - Global Analyses of Human Immune Variation Reveal Baseline Predictors of Postvaccination Responses (PMID: 24725414)"
author: "Mikkel Niklas Rasmussen"
date: "4/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Applying the the diagonal linear discriminant analysis (DLDA) using cell population frequencies on the CMI-PD dataset

```{r load libaries, echo=FALSE}

# Set directory to file location
setwd(dirname(rstudioapi::getSourceEditorContext()$path))

suppressMessages(library('data.table'))
suppressMessages(library('dplyr'))
suppressMessages(library('plyr'))
suppressMessages(library('tidyr'))
suppressMessages(library('Biobase'))
suppressMessages(library('DESeq2'))
suppressMessages(library('pROC'))
suppressMessages(library("ggplot2"))
suppressMessages(library("RCurl"))
suppressMessages(library("readr"))
suppressMessages(library("biomaRt"))
suppressMessages(library('klaR'))
suppressMessages(library("tibble"))
suppressMessages(library("readxl"))

# Load functions for evaluating gene signatures
source('../R/modelEvaluate.R')
source("../dlda/internals.r")

```

#### Load the CMI-PB datasets

```{r load data}

DATA_DIR_raw <- "../Data/raw"
DATA_DIR_full <- "../Data/processed/full"

# Info table (meta data)
meta.2020 <- fread(file.path(DATA_DIR_full, "meta.2020.pivoted.tsv"))
meta.2021 <- fread(file.path(DATA_DIR_full, "meta.2021.pivoted.tsv"))
specimen.table.2020 <- fread(file.path(DATA_DIR_raw, "2020LD_specimen.csv"))
specimen.table.2021 <- fread(file.path(DATA_DIR_raw, "2021BD_specimen.csv"))

# Antibody titer data
ab.titer.2020 <- fread(file.path(DATA_DIR_raw, "2020LD_ab_titer.csv"))
ab.titer.2021.day0 <- fread(file.path(DATA_DIR_raw, "2021BD_ab_titer.csv"))
ab.titer.2021.day14 <- fread(file.path(DATA_DIR_raw, "../2021_ab_titer_day14.csv"))

# Load cell population info table from the original study
cell.pop.map <- read_xlsx("Annotation_of_Cell_Subpopulations.xlsx",
                          range = "A2:I127") %>% 
   dplyr::select(-one_of("...6"))

# Cell type CMI-PB info table
cell.types <- fread(file.path(DATA_DIR_raw, "../cell_types.txt"))
cell.types.2020 <- cell.types %>% 
   filter(dataset == "2020_dataset")
cell.types.2021 <- cell.types %>% 
   filter(dataset == "2021_dataset")

# Cell frequnecy data
cytof.2020 <- fread(file.path(DATA_DIR_raw, "2020LD_live_cell_percentages.csv"))
cytof.2021 <- fread(file.path(DATA_DIR_raw, "2021BD_live_cell_percentages.csv"))

```

## Step 1: Preprocessing antibody titer

```{r}

#### Antibody titer data
# Normalized antibody titer data by baseline median
abt.2020.all <- ab.titer.2020 %>% 
   filter(antigen %in% c('PT', 'PRN', 'FHA', 'FIM2/3')) %>% 
   unite(isotype_antigen, c("isotype", "antigen"), sep = "-") %>% 
   inner_join(., specimen.table.2020 %>% dplyr::select(specimen_id, subject_id, 
                                                  planned_day_relative_to_boost), 
              by='specimen_id') %>% 
   filter(planned_day_relative_to_boost %in% c(0, 1, 3, 7, 14, 30)) %>% 
   dplyr::rename(timepoint = planned_day_relative_to_boost) %>% 
   tibble()

# Select the subject that have data points on all days
subjects.day0 <- unique(abt.2020.all[abt.2020.all$timepoint == 0, ]$subject_id)
subjects.day14 <- unique(abt.2020.all[abt.2020.all$timepoint == 14, ]$subject_id)
ids <- intersect(subjects.day0, subjects.day14)

# Subset antibody titer
abt.2020.subset <- abt.2020.all[abt.2020.all$subject_id %in% ids, ]
abt.2020.day0 <- abt.2020.subset[abt.2020.subset$timepoint == 0, ]
abt.2020.day14 <- abt.2020.subset[abt.2020.subset$timepoint == 14, ]

```

## Step 2: Preprocess CMI-PB 2020 cell population data

```{r}
#### Cell frequency data
# Merge cell frequnecy data with subject and day information
cytof.2020 <- cytof.2020 %>% 
   inner_join(., specimen.table.2020 %>% dplyr::select(specimen_id, subject_id, 
                                                  planned_day_relative_to_boost), 
              by='specimen_id') %>% 
   filter(planned_day_relative_to_boost %in% c(0, 1, 3, 7, 14)) %>% 
   dplyr::rename(timepoint = planned_day_relative_to_boost) %>% 
   tibble()

# Join cell type and gating information with cell frequency data
cytof.2020 <- cytof.2020 %>% 
   inner_join(., cell.types.2020, by="cell_type_name")

# Select day 0 (baseline) cell frequencies and transform with log10, as 
# described in the orignal study (all zero values were set to 0.01 before 
# transformation)
cytof.2020.day0 <- cytof.2020[cytof.2020$timepoint == 0, ]
cytof.2020.day0[cytof.2020.day0$percent_live_cell == 0, ]$percent_live_cell <- 0.01
cytof.2020.day0$percent_live_cell <- log10(cytof.2020.day0$percent_live_cell)

```

## Step 3: Manually selection of the cell populations in the CMI-PB 2020 to be used in the DLDA models

```{r}

# The top cell populations identified using the dlda_example.R provided by the
# authors
top2 <- c("ID36", "ID103")
top5 <- c("ID36", "ID103", "ID37", "ID52.T1", "ID96")

### Top 2 cell populations
for(cell.pop in cell.pop.map[cell.pop.map$`Population ID` %in% top2, ]$`Subset name`){
   print(cell.pop)
}

# Effector memory CD4+ T cells - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top2, ]$Markers[1]
print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2020[cell.types.2020$cell_type_name == 'TemCD4', ]$gating_definition

# Memory B cells IgD+ CD38+ (closest in CMI-PB is Bcells) - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top2, ]$Markers[2] 
print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2020[cell.types.2020$cell_type_name == 'ASCs (Plasmablasts)', ]$gating_definition

# The selected cell populations closest to the top 2 in the CMI-PB dataset
populations.top2.2020 <- c('TemCD4', 'ASCs (Plasmablasts)')

### Top 5 cell populations
for(cell.pop in cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$`Subset name`){
   print(cell.pop)
}

# Effector memory CD4+ T cells - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[1]

print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2020[cell.types.2020$cell_type_name == 'TemCD4', ]$gating_definition

# Central memory T helper cells - gating definitions: 
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[2]

print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2020[cell.types.2020$cell_type_name == "TcmCD4", ]$gating_definition

# T cytotoxic CD8+ CD27+ (closest in CMI-PB is CD8Tcells) - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[3]

print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2020[cell.types.2020$cell_type_name == "CD8Tcells", ]$gating_definition

# Memory B cells IgD- CD38+
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[4] 
print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2020[cell.types.2020$cell_type_name == 'ASCs (Plasmablasts)', ]$gating_definition

# Memory B cells IgD+ CD38+
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[5]
print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2020[cell.types.2020$cell_type_name == "Bcells", ]$gating_definition

# The selected cell populations closest to the top 5 in the CMI-PB dataset
populations.top5.2020 <- c("TemCD4", "TcmCD4", "CD8Tcells", "Bcells", 'ASCs (Plasmablasts)')


```

## Step 4: Evaluate the DLDA model using different number of cell populations on the CMI-PB 2020

```{r model evaluation}

DLDA.models <- readRDS("../dlda/DLDA_models.rds")

# Isotype-antigens which are considered the most important
selected.iso.ag <- c("IgG-PT", "IgG-FHA",   "IgG-PRN", "IgG1-PT",
                     "IgG1-FHA", "IgG4-PT", "IgG4-FHA")

# Calculate the class probabilities for the DLDA model using 2 cell populations
df.DLDA.top2.2020.pred <- mlModelPredict(cytof.data=cytof.2020.day0,
                                    populations=populations.top2.2020,
                                    study='Tsang_2014', 
                                    model=DLDA.models[[1]],
                                    filter.ab.ag=selected.iso.ag,
                                    post.ab.data=abt.2020.day14,
                                    pre.ab.data=abt.2020.day0)

# Evaluate DLDA model with 2 cell populations
Tsang.2014.DLDA.top2.2020.results <- evaluateModel(df.prediction=df.DLDA.top2.2020.pred, 
                                              direction = 'missing', N.perm = 1000,
                                              score.label='Class probability',
                                              add.to.title = " -  DLDA model with 2 cell populations",
                                              fn1='DLDApred_vs_day14_heatmap',
                                              fn2='DLDApred_vs_day14_correlation',
                                              fn3='DLDApred_vs_FC_heatmap',
                                              fn4='DLDApred_vs_FC_correlation',
                                              fn5='DLDApred_vs_day14_AUC',
                                              fn6='DLDApred_vs_day14_boxplot',
                                              fn7='DLDApred_vs_FC_AUC',
                                              fn8='DLDApred_vs_FC_boxplot',
                                              path = 'Results/2020/DLDA_top2/')

save(Tsang.2014.DLDA.top2.2020.results, 
     file = "Results/2020/DLDA_top2/Tsang.2014.DLDA.top2.2020.results.RData")


# Calculate the class probabilities for the DLDA model using 5 cell populations
df.DLDA.top5.2020.pred <- mlModelPredict(cytof.data=cytof.2020.day0,
                                         populations=populations.top5.2020,
                                         study='Tsang_2014', 
                                         model=DLDA.models[[2]],
                                         filter.ab.ag=selected.iso.ag,
                                         post.ab.data=abt.2020.day14,
                                         pre.ab.data=abt.2020.day0)


# Evaluate DLDA model with 5 cell populations
Tsang.2014.DLDA.top5.2020.results <- evaluateModel(df.prediction=df.DLDA.top5.2020.pred, 
                                                   direction = 'missing', N.perm = 1000,
                                                   score.label='Class probability',
                                                   add.to.title = " -  DLDA model with 5 cell populations",
                                                   fn1='DLDApred_vs_day14_heatmap',
                                                   fn2='DLDApred_vs_day14_correlation',
                                                   fn3='DLDApred_vs_FC_heatmap',
                                                   fn4='DLDApred_vs_FC_correlation',
                                                   fn5='DLDApred_vs_day14_AUC',
                                                   fn6='DLDApred_vs_day14_boxplot',
                                                   fn7='DLDApred_vs_FC_AUC',
                                                   fn8='DLDApred_vs_FC_boxplot',
                                                   path = 'Results/2020/DLDA_top5/')

save(Tsang.2014.DLDA.top5.2020.results, 
     file = "Results/2020/DLDA_top5/Tsang.2014.DLDA.top5.2020.results.RData")

```

## Step 5: Pre-process CMI-PB 2021 antibody titer data

```{r}

# Normalized antibody titer data by baseline median
ab.titer.day0.2021.all <- ab.titer.2021.day0 %>% 
   filter(antigen %in% c('PT', 'PRN', 'FHA', 'FIM2/3')) %>% 
   unite(isotype_antigen, c("isotype", "antigen"), sep = "-") %>% 
   inner_join(., specimen.table.2021 %>% dplyr::select(specimen_id, subject_id, 
                                                  planned_day_relative_to_boost), 
              by='specimen_id') %>% 
   dplyr::rename(timepoint = planned_day_relative_to_boost) %>% 
   tibble()

ab.titer.subject.ids <- unique(ab.titer.day0.2021.all$subject_id)
abt.2021.day0 <- ab.titer.day0.2021.all[ab.titer.day0.2021.all$timepoint == 0, ]
abt.2021.day0 <- abt.2021.day0 %>% 
   filter(isotype_antigen %in% selected.iso.ag)

# Change format of the day 14 antibody titers
abt.2021.day14 <- ab.titer.2021.day14 %>% 
   separate(task, into=c("task", "antigen"), sep = "\\) ") %>% 
   dplyr::mutate(isotype_antigen = recode(antigen, 
                                          `Pertussis Toxin` = "IgG-PT", 
                                          `FHA` = 'IgG-FHA', 
                                          `Pertactin` =  'IgG-PRN',
                                          `IgG1 - Pertussis toxin` = "IgG1-PT",
                                          `IgG1 - FHA` = "IgG1-FHA",
                                          `IgG4 - Pertussis toxin` = "IgG4-PT",
                                          `IgG4 - FHA` = "IgG4-FHA"))


```

Step 6: Preprocess CMI-PB 2021 cell population data

```{r}
#### Cell frequency data
# Merge cell frequnecy data with subject and day information
cytof.2021 <- cytof.2021 %>% 
   inner_join(., specimen.table.2021 %>% dplyr::select(specimen_id, subject_id, 
                                                  planned_day_relative_to_boost), 
              by='specimen_id') %>% 
   filter(planned_day_relative_to_boost %in% c(0, 1, 3, 7, 14)) %>% 
   dplyr::rename(timepoint = planned_day_relative_to_boost) %>% 
   tibble()

# Subset cell frequency data based on subjects with antibody titers
cytof.2021 <- cytof.2021[cytof.2021$subject_id %in% ab.titer.subject.ids, ]

# Join cell type and gating information with cell frequency data
cytof.2021 <- cytof.2021 %>% 
   inner_join(., cell.types.2021, by="cell_type_name")

# Select day 0 (baseline) cell frequencies and transform with log10, as 
# described in the original study (all zero values were set to 0.01 before 
# transformation)
cytof.2021.day0 <- cytof.2021[cytof.2021$timepoint == 0, ]
cytof.2021.day0[is.na(cytof.2021.day0$percent_live_cell), ]$percent_live_cell <- 0.01
cytof.2021.day0[cytof.2021.day0$percent_live_cell == 0, ]$percent_live_cell <- 0.01
cytof.2021.day0$percent_live_cell <- log10(cytof.2021.day0$percent_live_cell)

```

Step 7: Manually select cell populations from the CMI-PB 2021 dataset

```{r}

# The top cell populations identified using the dlda_example.R provided by the
# authors
top2 <- c("ID36", "ID103")
top5 <- c("ID36", "ID103", "ID37", "ID52.T1", "ID96")

### Top 2 cell populations
for(cell.pop in cell.pop.map[cell.pop.map$`Population ID` %in% top2, ]$`Subset name`){
   print(cell.pop)
}

# Effector memory CD4+ T cells - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top2, ]$Markers[1]
print("Closest cell population in CMI-PB 2021 dataset - gatinge definition:")
cell.types.2021[cell.types.2021$cell_type_name == 'TemCD4', ]$gating_definition

# Memory B cells IgD+ CD38+ (closest in CMI-PB is Bcells) - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top2, ]$Markers[2] 
print("Closest cell population in CMI-PB 2021 dataset - gatinge definition:")
cell.types.2021[cell.types.2021$cell_type_name == 'Naive B cells', ]$gating_definition

# The selected cell populations closest to the top 2 in the CMI-PB dataset
populations.top2.2021 <- c('TemCD4', 'Naive B cells')

### Top 5 cell populations
for(cell.pop in cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$`Subset name`){
   print(cell.pop)
}

# Effector memory CD4+ T cells - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[1]

print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2021[cell.types.2021$cell_type_name == 'TemCD4', ]$gating_definition

# Central memory T helper cells - gating definitions: 
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[2]

print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2021[cell.types.2021$cell_type_name == "TcmCD4", ]$gating_definition

# T cytotoxic CD8+ CD27+ (closest in CMI-PB is CD8Tcells) - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[3]

print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2021[cell.types.2021$cell_type_name == "CD8Tcells", ]$gating_definition

# Memory B cells IgD- CD38+ - gating definitions:
print("Original study - gating definition: ")
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[4] 

print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2021[cell.types.2021$cell_type_name == 'Memory B cells', ]$gating_definition

# Memory B cells IgD+ CD38+ - gating definitions:
cell.pop.map[cell.pop.map$`Population ID` %in% top5, ]$Markers[5]

print("Closest cell population in CMI-PB 2020 dataset - gatinge definition:")
cell.types.2021[cell.types.2021$cell_type_name == 'Naive B cells', ]$gating_definition

# The selected cell populations closest to the top 5 in the CMI-PB dataset
populations.top5.2021 <- c("TemCD4", "TcmCD4", "CD8Tcells", 'Memory B cells', 'Naive B cells')

```

Step 8: Make prediction for CMI-PB 2021 dataset

```{r}

# Calculate the class probabilities for the DLDA model using 2 cell populations
df.DLDA.top2.2021.pred <- mlModelPredict(cytof.data=cytof.2021.day0,
                                         populations=populations.top2.2021,
                                         study='Tsang_2014', 
                                         model=DLDA.models[[1]])

# Calculate the class probabilities for the DLDA model using 5 cell populations
df.DLDA.top5.2021.pred <- mlModelPredict(cytof.data=cytof.2021.day0,
                                  populations=populations.top5.2021,
                                  study='Tsang_2014', 
                                  model=DLDA.models[[2]])


# Read submisson format
submission.format <- fread("../../Data/2021_cmipb_challenge/2021_submission_format.tsv") %>% 
   dplyr::rename(subject_id = Subject_ID)

# DLDA model using 2 cell populations
Tsang.2014.DLDA.top2.2021.pred <- submission.format %>% 
   left_join(., df.DLDA.top2.2021.pred, by="subject_id")
Tsang.2014.DLDA.top2.2021.pred[, c(5:11)] <- Tsang.2014.DLDA.top2.2021.pred$rank.predictor
Tsang.2014.DLDA.top2.2021.pred <- Tsang.2014.DLDA.top2.2021.pred[,c(1:18)] %>% 
   dplyr::rename(Subject_ID = subject_id)
write.table(Tsang.2014.DLDA.top2.2021.pred, 
            file = "Results/2021_predictions/tsang_2014_DLDA_top2_CMI-PB_2021.tsv",
            quote=FALSE, sep='\t', row.names = FALSE)


# DLDA model using 5 cell populations
Tsang.2014.DLDA.top5.2021.pred <- submission.format %>% 
   left_join(., df.DLDA.top5.2021.pred, by="subject_id")
Tsang.2014.DLDA.top5.2021.pred[, c(5:11)] <- Tsang.2014.DLDA.top5.2021.pred$rank.predictor
Tsang.2014.DLDA.top5.2021.pred <- Tsang.2014.DLDA.top5.2021.pred[,c(1:18)] %>% 
   dplyr::rename(Subject_ID = subject_id)
write.table(Tsang.2014.DLDA.top5.2021.pred, 
            file = "Results/2021_predictions/tsang_2014_DLDA_top5_CMI-PB_2021.tsv",
            quote=FALSE, sep='\t', row.names = FALSE)

```

## Step 9: Evaluate the DLDA model using different number of cell populations on the CMI-PB 2021 dataset

```{r model evaluation}

# Calculate the class probabilities for the DLDA model using 2 cell populations
df.DLDA.top2.2021.pred <- mlModelPredict(cytof.data=cytof.2021.day0,
                                         populations=populations.top2.2021,
                                         study='Tsang_2014', 
                                         model=DLDA.models[[1]],
                                         filter.ab.ag=selected.iso.ag,
                                         post.ab.data=abt.2021.day14,
                                         pre.ab.data=abt.2021.day0)

# Evaluate DLDA model with 2 cell populations
Tsang.2014.DLDA.top2.2021.results <- evaluateModel(df.prediction=df.DLDA.top2.2021.pred, 
                                                   direction = 'missing', N.perm = 1000,
                                                   score.label='Class probability',
                                                   add.to.title = " -  DLDA model with 2 cell populations",
                                                   fn1='DLDApred_vs_day14_heatmap',
                                                   fn2='DLDApred_vs_day14_correlation',
                                                   fn3='DLDApred_vs_FC_heatmap',
                                                   fn4='DLDApred_vs_FC_correlation',
                                                   fn5='DLDApred_vs_day14_AUC',
                                                   fn6='DLDApred_vs_day14_boxplot',
                                                   fn7='DLDApred_vs_FC_AUC',
                                                   fn8='DLDApred_vs_FC_boxplot',
                                                   path = 'Results/2021/DLDA_top2_2021_populations/')

save(Tsang.2014.DLDA.top2.2021.results, 
     file = "Results/2021/DLDA_top2_2021_populations/Tsang.2014.DLDA.top2.2021.results.RData")


# Calculate the class probabilities for the DLDA model using 5 cell populations
df.DLDA.top5.2021.pred <- mlModelPredict(cytof.data=cytof.2021.day0,
                                         populations=populations.top5.2021,
                                         study='Tsang_2014', 
                                         model=DLDA.models[[2]],
                                         filter.ab.ag=selected.iso.ag,
                                         post.ab.data=abt.2021.day14,
                                         pre.ab.data=abt.2021.day0)


# Evaluate DLDA model with 5 cell populations
Tsang.2014.DLDA.top5.2021.results <- evaluateModel(df.prediction=df.DLDA.top5.2021.pred, 
                                                   direction = 'missing', N.perm = 1000,
                                                   score.label='Class probability',
                                                   add.to.title = " -  DLDA model with 5 cell populations",
                                                   fn1='DLDApred_vs_day14_heatmap',
                                                   fn2='DLDApred_vs_day14_correlation',
                                                   fn3='DLDApred_vs_FC_heatmap',
                                                   fn4='DLDApred_vs_FC_correlation',
                                                   fn5='DLDApred_vs_day14_AUC',
                                                   fn6='DLDApred_vs_day14_boxplot',
                                                   fn7='DLDApred_vs_FC_AUC',
                                                   fn8='DLDApred_vs_FC_boxplot',
                                                   path = 'Results/2021/DLDA_top5_2021_populations/')

save(Tsang.2014.DLDA.top5.2021.results, 
     file = "Results/2021/DLDA_top5_2021_populations/Tsang.2014.DLDA.top5.2021.results.RData")

```

## Step 9: Evaluate the DLDA model using different number of cell populations on the CMI-PB 2021 dataset

```{r model evaluation}

# Calculate the class probabilities for the DLDA model using 2 cell populations
df.DLDA.top2.2021.pred <- mlModelPredict(cytof.data=cytof.2021.day0,
                                         populations=populations.top2.2020,
                                         study='Tsang_2014', 
                                         model=DLDA.models[[1]],
                                         filter.ab.ag=selected.iso.ag,
                                         post.ab.data=abt.2021.day14,
                                         pre.ab.data=abt.2021.day0)

# Evaluate DLDA model with 2 cell populations
Tsang.2014.DLDA.top2.2021.results <- evaluateModel(df.prediction=df.DLDA.top2.2021.pred, 
                                                   direction = 'missing', N.perm = 1000,
                                                   score.label='Class probability',
                                                   add.to.title = " -  DLDA model with 2 cell populations",
                                                   fn1='DLDApred_vs_day14_heatmap',
                                                   fn2='DLDApred_vs_day14_correlation',
                                                   fn3='DLDApred_vs_FC_heatmap',
                                                   fn4='DLDApred_vs_FC_correlation',
                                                   fn5='DLDApred_vs_day14_AUC',
                                                   fn6='DLDApred_vs_day14_boxplot',
                                                   fn7='DLDApred_vs_FC_AUC',
                                                   fn8='DLDApred_vs_FC_boxplot',
                                                   path = 'Results/2021/DLDA_top2_2020_populations/')

save(Tsang.2014.DLDA.top2.2021.results, 
     file = "Results/2021/DLDA_top2_2020_populations/Tsang.2014.DLDA.top2.2021.results.RData")


# Calculate the class probabilities for the DLDA model using 5 cell populations
df.DLDA.top5.2021.pred <- mlModelPredict(cytof.data=cytof.2021.day0,
                                         populations=populations.top5.2020,
                                         study='Tsang_2014', 
                                         model=DLDA.models[[2]],
                                         filter.ab.ag=selected.iso.ag,
                                         post.ab.data=abt.2021.day14,
                                         pre.ab.data=abt.2021.day0)


# Evaluate DLDA model with 5 cell populations
Tsang.2014.DLDA.top5.2021.results <- evaluateModel(df.prediction=df.DLDA.top5.2021.pred, 
                                                   direction = 'missing', N.perm = 1000,
                                                   score.label='Class probability',
                                                   add.to.title = " -  DLDA model with 5 cell populations",
                                                   fn1='DLDApred_vs_day14_heatmap',
                                                   fn2='DLDApred_vs_day14_correlation',
                                                   fn3='DLDApred_vs_FC_heatmap',
                                                   fn4='DLDApred_vs_FC_correlation',
                                                   fn5='DLDApred_vs_day14_AUC',
                                                   fn6='DLDApred_vs_day14_boxplot',
                                                   fn7='DLDApred_vs_FC_AUC',
                                                   fn8='DLDApred_vs_FC_boxplot',
                                                   path = 'Results/2021/DLDA_top5_2020_populations/')

save(Tsang.2014.DLDA.top5.2021.results, 
     file = "Results/2021/DLDA_top5_2020_populations/Tsang.2014.DLDA.top5.2021.results.RData")

```
